### 2016-06-06 AnalysisCode/burden_mC.Rmd
first commit

---
title: "burden_mC"
author: "Tim Stuart"
date: "6 June 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import}
library(dplyr)
```

```{r load_data}
mC <- read.table("../data/c_dmr_mc.tsv", header = T)
TE_DMR <- read.table("../data/")
```

```{r functions}
gatherData <- function(insertions, ranks) {
  data <- vector()
  for(i in 1:nrow(insertions)){
    row <- insertions[i,]
    gene <- as.character(row[[1]])
    accessions <- unlist(strsplit(gsub("-", "_", as.character(row[[3]])), ","))
    accessions <- intersect(accessions, colnames(ranks))
    if(length(accessions) > 0){
      for(acc in accessions){
        data <- c(data, as.numeric(ranks[gene,acc]))
      }
    }
  }
  return(data)
}

permuteLabels <- function(ranks, insertions) {
  data <- vector()
  all_names <- colnames(ranks)
  for(i in 1:nrow(insertions)){
    row <- insertions[i,]
    gene <- as.character(row[[1]])
    accessions <- unlist(strsplit(gsub("-", "_", as.character(row[[3]])), ","))
    accessions <- intersect(accessions, colnames(ranks))
    if(length(accessions) > 0){
      rand <- sample(all_names, length(accessions), replace = FALSE)
      for(acc in rand){
        data <- c(data, as.numeric(ranks[gene,acc]))
      }
    }
  }
  return(data)
}
```