### 2016-08-11 Paper/paper.tex
Add paper

\documentclass[12pt]{article}
\usepackage[labelfont=bf]{caption}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{newfloat}
\usepackage[flushleft]{threeparttable}
\usepackage{parskip}
\usepackage[
  backend=bibtex,
  style=numeric,
  sorting=none,
  doi=true,
  url=false,
  citestyle=numeric-comp
]{biblatex}

\AtEveryBibitem{%
  \clearlist{language}%
  }
\setlength\bibitemsep{\baselineskip}

\bibliography{references.bib}
\usepackage{doi}
\usepackage{authblk}

\DeclareFloatingEnvironment[name={Figure}]{suppfigure}
\usepackage{amssymb,amsmath}
\usepackage[margin=2cm]{geometry}
\graphicspath{ {../Plots/FinalFigures/} }
\usepackage{lineno}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true,
            colorlinks=true,
            citecolor=blue
}
\urlstyle{same}  % don't use monospace font for urls
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\date{}

\begin{document}

\title{Population scale mapping of transposable element diversity reveals links to gene regulation and epigenomic variation}

\author[1]{Tim Stuart}
\author[2]{Steven R. Eichten}
\author[1]{Jonathan Cahn}
\author[1]{Yuliya Karpievitch}
\author[2]{Justin Borevitz}
\author[1]{Ryan Lister}
\affil[1]{ARC Centre of Excellence in Plant Energy Biology, The University of
Western Australia, Perth, Australia}
\affil[2]{ARC Centre of Excellence in Plant Energy Biology, The Australian
National University, Canberra, Australia}

\renewcommand\Authands{ and }

\maketitle

Corresponding author: Ryan Lister \href{mailto:ryan.lister@uwa.edu.au}{ryan.lister@uwa.edu.au} \\

\textbf{Author ORCID IDs:}

0000-0002-3044-0897 (TS)

0000-0003-2268-395X (SRE)

0000-0002-5006-741X (JC)

0000-0001-6637-7239 (RL)

\linenumbers

\section{Abstract}

Variation in the presence or absence of transposable elements (TEs) is a
major source of genetic variation between individuals. Here, we
identified 23,095 TE presence/absence variants between 216 wild
Arabidopsis accessions. Most variants are rare, and we find a burden of
rare variants associated with extremes of both gene expression and DNA
methylation levels within the population. Of the common alleles
identified, two thirds were not in linkage disequilibrium with nearby
SNPs, implicating these TE variants as a source of novel genetic
diversity. Nearly 200 common TE variants were associated with
significantly altered expression of nearby genes, with potential
functional consequences including decreased pathogen resistance. A major
fraction of inter-accession DNA methylation differences were associated
with nearby TE insertion variants, with DNA, indicating an important
role of TE variants in facilitating epigenomic variation.

\section{Introduction}

Transposable elements (TEs) are mobile genetic elements present in
nearly all studied organisms, and make up a large fraction of most
eukaryotic genomes. The two classes of TEs are retrotransposons (class
I elements), which transpose via an RNA intermediate requiring a
reverse transcription reaction, and DNA transposons (class II
elements), which transpose via either a cut-paste or, in the case of
Helitrons, a rolling circle mechanism with no RNA intermediate
\cite{Wicker:2007en}. TE activity poses mutagenic potential as a TE
insertion may disrupt essential regions of the genome. Consequently,
safeguards have evolved in order to suppress this activity. These
safeguards include epigenetic transcriptional silencing mechanisms,
chiefly involving the methylation of cytosine nucleotides (DNA
methylation) to produce 5-methylcytosine (mC), a mark that can signal
transcriptional silencing of the methylated locus. In
\emph{Arabidopsis thaliana }(Arabidopsis), DNA methylation occurs in
three DNA sequence contexts: mCG, mCHG, and mCHH, where H is any base
but G. Establishment of DNA methylation marks can be carried out by
two distinct pathways -- the RNA-directed DNA methylation pathway
guided by 24 nucleotide (nt) small RNAs (smRNAs), and the DDM1/CMT2
pathway \cite{Zemach:2013dj, Matzke:2014ek}. A major function of DNA
methylation in Arabidopsis is in the transcriptional silencing of TEs.
Loss of DNA methylation due to mutations in genes essential for DNA
methylation establishment and maintenance leads to expression of
previously silent TEs, and in some cases transposition
\cite{Mirouze:2009km, Miura:2001eg, Saze:2003du, Lippman:2004cm,
  Jeddeloh:1999gu, Zemach:2013dj}.

TEs are thought to play an important role in evolution, not only because
of the disruptive potential of their transposition. The release of
transcriptional and post-transcriptional silencing of TEs can lead to
bursts of TE activity, quickly generating new genetic diversity
\cite{Vitte:2014he}. TEs may carry regulatory information such as
promoters and transcription factor binding sites, and their mobilization
may lead to the creation or expansion of gene regulatory networks
\cite{Henaff:2014hg, Bolger:2014bn, Ito:2011dga, Makarevitch:2015ho}.
Furthermore, the transposase enzymes required and encoded by TEs have
frequently been domesticated and repurposed as endogenous proteins, such
as the \emph{DAYSLEEPER} gene in Arabidopsis, derived from a hAT
transposase enzyme \cite{Bundock:2005gp}. Clearly, the activity of TEs
can have widespread and unpredictable effects on the host genome.
However, the identification of TE presence/absence variants in genomes
has remained difficult to date. It is challenging to identify the
structural changes in the genome caused by TE mobilization using current
short-read sequencing technologies as these reads are typically mapped
to a reference genome, which has the effect of masking structural
changes that may be present. However, in terms of the number of base
pairs affected, a large fraction of genetic differences between
Arabidopsis accessions appear to be due to variation in TE content
\cite{Cao:2011cf, Quadrana:2016bi}. Therefore identification of TE
variants is essential in order to develop a more comprehensive
understanding of the genetic variation that exists between genomes, and
of the consequences of TE movement on genome and cellular function.

The tools developed previously for identification of novel TE insertion
sites have several limitations. They either require a library of active
TE sequences, cannot identify TE absence variants, or are not designed
with population studies in mind, or suffer from a high degree of
false-negatives
\cite{Thung:2014ir, Robb:2013gw, Henaff:2015dl, Quadrana:2016bi}. In
order to accurately map the locations of TE presence/absence variants
with respect to a reference genome, we have developed a novel algorithm,
TEPID (Transposable Element Polymorphism IDentification), which is
designed for population studies. We tested our algorithm using both
simulated and real Arabidopsis sequencing data, finding that TEPID is
able to accurately identify TE presence/absence variants with respect to
the Col-0 reference genome. We applied our TE analysis method to
existing genome resequencing data for 216 different wild Arabidopsis
accessions, and identified widespread TE variation amongst these
accessions \cite{Schmitz:2013iu}.

\section{Results}

\subsection{Computational identification of TE presence/absence variation}

We developed TEPID, an analysis pipeline capable of detecting TE
presence/absence variants from paired end DNA sequencing data. TEPID
integrates split and discordant read mapping information, read mapping
quality, sequencing breakpoints, as well as local variations in
sequencing coverage to identify novel TE presence/absence variants with
respect to a reference TE annotation (Figure \ref{fig1}; see methods). This
typically takes 5-10 minutes using Arabidopsis sequencing data at 20-40x
coverage (excluding the read mapping step). After TE variant discovery
has been performed, TEPID then includes a second refinement step
designed for population studies. This examines each region of the genome
where there was a TE insertion identified in any member of the group,
and checks for evidence of this insertion in all other members of the
group. In this way, TEPID leverages TE variant information for a group
of related samples to correct false negative calls within the group.
Testing of TEPID using simulated TE variants in the Arabidopsis genome
showed that it was able to reliably detect simulated TE variants at
sequencing coverage levels commonly used in genomics studies (Figure 1 -
figure supplement \ref{fig1s1}).

In order to further assess the sensitivity and specificity of TE variant
discovery using TEPID, we identified TE variants in the \emph{Landsberg}
\emph{erecta} (L\emph{er}) accession, and compared these with the
L\emph{er }genome assembly created using long PacBio sequencing reads
\cite{Chin:2013iw}. Previously published 100 bp paired-end L\emph{er}
genome resequencing reads \cite{Schneeberger:2011ft} were first analyzed
using TEPID, enabling identification of 446 TE insertions (Figure 1 -
source data 1) and 758 TE absence variants (Figure 1 - source data 2)
with respect to the Col-0 reference TE annotation. Reads providing
evidence for these variants were then mapped to the L\emph{er} reference
genome that was generated by \emph{de novo} assembly using Pacific
Biosciences P5-C3 chemistry with a 20 kb insert library
\cite{Chin:2013iw}, using the same alignment parameters as was used to
map reads to the Col-0 reference genome. This resulted in 98.7\% of
reads being aligned concordantly to the L\emph{er} reference, whereas
100\% aligned discordantly or as split reads to the Col-0 reference
genome (Table \ref{table1}). To find whether reads mapped to homologous regions in
both the Col-0 and L\emph{er} reference genomes, we conducted a blast
search \cite{Camacho:2009fc} using the DNA sequence between read pair
mapping locations in the L\emph{er }genome against the Col-0 genome, and
found the top blast result for 80\% of reads providing evidence for TE
insertions, and 89\% of reads providing evidence for TE absence variants
in L\emph{er}, to be located within 200 bp of the TE variant
reported by TEPID. We conclude that reads providing evidence for TE
variants map discordantly or as split reads when mapped to the Col-0
reference genome, but map concordantly to homologous regions of the
L\emph{er} \emph{de novo} assembled reference genome, indicating that
structural variation is present at the sites identified by TEPID, and
that this is resolved in the \emph{de novo} assembled genome.

To estimate the rate of false negative TE absence calls made using
TEPID, we compared our L\emph{er} TE absence calls to the set of TE
absences in L\emph{er} genome identified previously by aligning
full-length Col-0 TEs to the Ler reference using BLAT
\cite{Quadrana:2016bi}. We found that 89.6\% (173/193) of these TE
absences were also identified using TEPID, indicating a false negative
rate of around 10\% for TE absence calls. To determine the rate of false
negative TE insertion calls, we ran TEPID using 90 bp paired-end Col-0
reads (Col-0 control samples from \cite{Jiang:2014ih}), aligning reads
to the L\emph{er} PacBio assembly. As TEPID requires a high-quality TE
annotation to discover TE variants, which is not available for the
L\emph{er} assembly, we simply looked for discordant and split read
evidence at the known Col-0-specific TE insertion sites
\cite{Quadrana:2016bi}, and found evidence reaching the TEPID threshold
for a TE insertion call to be made at 89.6\% (173/193) of these sites,
indicating a false negative rate of 10\%. However, we note that this
estimate does not take into account the TEPID refinement step used on
large populations, and so the false negative rate for samples analyzed
in the population from Schmitz et al. (2013) is likely to be slightly
lower than this estimate, as each accession gained on average 4\% more
insertion calls following this refinement step (Figure 2 - figure
supplement \hyperref[fig2s1]{1}).

\subsection{Abundant TE positional variation among natural Arabidopsis
populations}

We used TEPID to analyze previously published 100 bp paired-end genome
resequencing data for 216 different Arabidopsis accessions
\cite{Schmitz:2013iu}, and identified 15,007 TE insertions and 8,088 TE
absence variants relative to the Col-0 reference accession, totalling
23,095 unique TE variants. In most accessions we identified 300-500 TE
insertions (mean = 378) and 1,000-1,500 TE absence variants (mean =
1,279), the majority of which were shared by two or more accessions
(Figure 2 - figure supplement \hyperref[fig2s2]{2}). PCR validations were performed for a
random subset of 10 insertions and 10 absence variants in 14 accessions
(totalling 280 validations), and confirmed the high accuracy of TE
variant discovery using the TEPID package, with results similar to that
observed using simulated data and the L\emph{er} genome analysis (Figure 2 -
figure supplement \hyperref[fig2s3]{3}). The number of TE insertions identified was
positively correlated with sequencing depth of coverage, while the
number of TE absence variants identified had no correlation with
sequencing coverage (Figure 2 - figure supplement \hyperref[fig2s4]{4}A, B), indicating
that the sensitivity of TE absence calls is not limited by sequencing
depth, while TE insertion calls benefitted from high sequencing depth.
However, accessions with low coverage gained more TE insertion calls
during the TEPID refinement step (Figure 2 - figure supplement \hyperref[fig2s4]{4}C),
indicating that these false negatives were effectively reduced by
leveraging TE variant information for the whole population.

As TE insertion and TE absence calls represent an arbitrary comparison
to the Col-0 reference genome, we sought to remove these arbitrary
comparisons and classify each variant as a new TE insertion or true
deletion of an ancestral TE in the population. To do this, we examined
the minor allele frequency (MAF) of each variant in the population,
under the expectation that the minor allele is the derived allele. We
re-classified common TE absences relative to Col-0 as rare TE insertions
in Col-0, and common TE insertions relative to Col-0 as true TE
deletions in Col-0. Cases where the TE variant was at an high MAF
(\textgreater{}20\%) were assigned NA calls, as it could not be
determined if these were cases where the variant was most likely to be
due to a true TE deletion or due to a new TE insertion. While these
classifications are not definitive, as there will be rare cases where a
true TE deletion has spread through the population and becomes the
common allele, we will correctly classify most TE variants. Overall, we
found 72.3\% of the TE absence variants identified with respect to the
Col-0 reference genome were likely due to a true TE deletion in these
accessions, while 4.8\% were due to new insertions in Col-0 not shared
by others in the population (Table \ref{table2}). Overall, we identified 15,077 new
TE insertions, 5,856 true TE deletions, and 2,162 TE variants at a high
MAF that were unable to be classified as an insertion or deletion.

TE insertions and deletions were distributed throughout chromosome 1 in
a pattern that was similar to the distribution of all Col-0 TEs (Figure
\ref{fig2}A). TE deletions and common TE variants were found in similar
chromosomal regions, as deletion variants represent the rare loss of
common variants. TE deletions and common variants were more highly
enriched in the pericentromeric regions than rare variants or TE
insertions. The distribution of rare TE variants and TE insertions was
somewhat similar to that observed for regions of the genome previously
identified as being differentially methylated in all DNA methylation
contexts (mCG, mCHG, mCHH) between the wild accessions (population
C-DMRs), while population CG-DMRs, differentially methylated in the mCG
context, less frequently overlapped with all types of TE variants
identified \cite{Schmitz:2013iu}. Furthermore, TE variants were depleted
within genes and DNase I hypersensitivity sites \cite{Sullivan:2014ii},
while they were enriched in gene flanking regions and within other
annotated TEs or pseudogenes (Figure \ref{fig2}B). We found that TE deletions and
common TE variants were enriched within the set of TE variants found in
gene bodies (Figure \ref{fig2}C, D). We did not find any significant enrichment
of TE variants within the \emph{KNOT ENGAGED ELEMENT (KEE)} regions,
previously identified as regions that may act as a ``TE sink''
\cite{Grob:2014kg} (Figure 2 - figure supplement \hyperref[fig2s5]{5}). This may indicate
that these regions do not act as a ``TE sink'' as has been previously
proposed, or that the ``TE sink'' activity is restricted to very recent
insertions, as the insertions we analysed in this population were likely
older than those used in the study by Grob et al.

Among the identified TE variants, several TE superfamilies were over- or
under-represented compared to the number expected by chance given the
overall genomic frequency of different TE types (Figure \ref{fig2}E). In
particular, both TE insertions and deletions in the RC/Helitron
superfamily were less numerous than expected, with an 11.5\% depletion
of RC/Helitron elements in the set of TE variants. In contrast, TEs
belonging to the LTR/Gypsy superfamily were more frequently deleted than
expected, with a 17\% enrichment in the set of TE deletions. This was
unlikely to be due to a differing ability of our detection methods to
identify TE variants of different lengths, as the TE variants identified
had a similar distribution of lengths as all Arabidopsis TEs annotated
in the Col-0 reference genome (Figure 2 - figure supplement \hyperref[fig2s6]{6}). These
enrichments suggest that the RC/Helitron TEs have been relatively
dormant in recent evolutionary history, while the LTR/Gypsy, which are
highly enriched in the pericentromeric regions, are frequently lost from
the Arabidopsis genome. At the family level, we observed similar
patterns of TE variant enrichment or depletion (Figure 2 - figure
supplement \hyperref[fig2s7]{7}; source data 5).

We further examined Arabidopsis (Col-0) DNA sequencing data from a
transgenerational stress experiment to investigate the possible minimum
number of generations required for TE variants to arise
\cite{Jiang:2014ih}. We identified a single potential TE insertion in a
sample following 10 generations of single-seed descent under high
salinity stress conditions, and no TE variants in the control
single-seed descent set. However, without experimental validation it
remains unclear if this represents a true variant. Therefore, we
conclude that TE variants may arise at a rate less than 1 insertion in
30 generations under laboratory conditions. Further experimental work
will be required to precisely determine the rate of transposition in
Arabidopsis.

\subsection{Relationship between TE variants and single nucleotide
polymorphisms}

Although thousands of TE variants were identified, they may be linked
(i.e. `tagged') by the previously identified single nucleotide
polymorphisms (SNPs), or unlinked from SNPs across the accessions. We
tested how often common TE variants (\textgreater{}3 \% MAF; 7
accessions) were linked to adjacent SNPs to determine when they would
represent a previously unassessed source of genetic variation between
accessions. SNPs that were previously identified between the accessions
\cite{Schmitz:2013iu} were compared to the presence/absence of
individual TE variants. For the testable TE variants in the population,
the nearest 300 flanking SNPs upstream and downstream of the TE variant
site were analyzed for local linkage disequilibrium (LD, r\textsuperscript{2}; see
methods). TE variants were classified as being either `low', `mid', or
`high' LD variants by comparing ranked r\textsuperscript{2} values of TE variant to SNPs
against the median ranked r\textsuperscript{2} value for all between SNP comparisons
(SNP-SNP) to account for regional variation in the extent of SNP-SNP LD
(Figure \ref{fig3}A, B) due to recombination rate variation or selection
\cite{Horton:2012jh}. The majority (61\%) of testable TE variants had
low LD with nearby SNPs, and represent a source of genetic diversity not
previously assessed by current SNP-based genotype calling methods
(Figure \ref{fig3}C). 29\% of TE variants displayed high levels of LD and are
tagged by nearby SNPs, while only 10\% had intermediate levels of LD. We
observed a positive correlation between TE variant MAF and LD state,
with variants of a high minor allele frequency more often classified as
high-LD (Figure \ref{fig3}D). While the proportion of TE variants classified as
high, mid, or low-LD was mostly the same for both TE insertions and TE
deletions, TE variants with a high MAF (\textgreater{}20\%) that were
unable to be classified as either true deletions or as new insertions
had a much higher proportion of high-LD variants (Figure \ref{fig3}E). This was
consistent with the observation that the more common alleles were more
often in a high-LD state. TE variants displayed a similar distribution
over chromosome one regardless of linkage classification (Figure 3 -
figure supplement \hyperref[fig3s1]{1}). Overall, this analysis revealed an abundance of
previously unexplored genetic variation that exists amongst Arabidopsis
accessions caused by the presence or absence of TEs, and illustrates the
importance of identifying TE variants alongside other genetic diversity
such as SNPs.

\subsection{TE variants affect gene expression}

To determine whether the TE variants identified affected nearby gene
expression, we compared the steady state transcript abundance within
mature leaf tissue between accessions with and without TE insertions
or deletions, for genes with TE variants located in the 2 kb gene
upstream region, 5' UTR, exon, intron, 3' UTR or 2 kb downstream
region (Figure \ref{fig4}A). While the steady state transcript
abundance of most genes appeared to be unaffected by the presence of a
TE, 196 genes displayed significant differences in transcript
abundance linked with the presence of a TE variant, indicating a role
for these variants in the local regulation of gene expression (1\%
false discovery rate; greater than 2-fold change in transcript
abundance; Figure \ref{fig4}A, Figure 4 - source data 1). We did
not find any functional category enrichments in this set of
differentially expressed genes. As rare TE variants with a MAF less
than 3\% may also be associated with a difference in transcript
abundance, but were unable to be statistically tested due to their
rarity, we performed a burden test for enrichment of rare variants in
the extremes of expression \cite{Zhao:2016gc}. Briefly, this method
counts the frequency of rare variants within each gene expression rank
in the population, and aggregates this information over the entire
population to determine whether an enrichment of rare variants exists
within any gene expression rank. We found a strong enrichment for gene
expression extremes for TE variants in the all gene features tested
(Figure \ref{fig4}B). While TE variants in gene upstream regions
showed a strong enrichment of both high and low gene expression ranks,
TE variants in exons or gene downstream regions seemed to have a
stronger enrichment for low expression ranks than high
ranks. Randomization of the accession names removed these enrichments
completely (Figure 4 - figure supplement \hyperref[fig4s1]{1}), and
there was little difference between TE insertions and TE deletions in
the gene expression rank enrichments found (Figure 4 - figure
supplement \hyperref[fig4s2]{2}). This rare variant analysis further
indicated that TE variants may alter the transcript abundance of
nearby genes.

As both increases and decreases in transcript abundance of nearby
genes were observed for TE variants within each gene feature, it
appears to be difficult to predict the impact a TE variant may have on
nearby gene expression. Furthermore, gene-level transcript abundance
measurements may fail to identify the potential positional effect of
TE variants upon transcription. To more closely examine changes in
transcript abundance associated with TE variants among the accessions,
we inspected a subset of TE variant sites and identified TE variants
that appear to have an impact on transcriptional patterns beyond
changes in total transcript abundance from a nearby gene. For example,
the presence of a TE insertion within an exon of \emph{AtRLP18
}(AT2G15040) was associated with truncation of the transcripts at the
TE insertion site in accessions possessing the TE variant, as well as
silencing of a downstream gene encoding a leucine-rich repeat protein
(AT2G15042) (Figure \ref{fig5}A, B). Both genes had significantly
lower transcript abundance in accessions containing the TE insertion
(p \textless{} $5.8\times10^{-10}$, Mann-Whitney U
test). \emph{AtRLP18 }is reported to be involved in bacterial
resistance, with the disruption of this gene by T-DNA insertion
mediated mutagenesis resulting in increased susceptibility to the
bacterial plant pathogen \emph{Pseudomonas syringae}
\cite{Wang:2008km}. We examined pathogen resistance phenotype data
\cite{Aranzana:2005cq}, and found that accessions containing the TE
insertion in the \emph{AtRLP18} exon were more often sensitive to
infection by \emph{Pseudomonas syringae} transformed with
\emph{avrPpH3} genes (Figure \ref{fig5}C). This suggests that the
accessions containing this TE insertion within \emph{AtRLP18} may have
an increased susceptibility to certain bacterial pathogens.

We also observed some TE variants associated with increased expression
of nearby genes. For example, the presence of a TE within the upstream
region of a gene encoding a pentatricopeptide repeat (PPR) protein
(AT2G01360) was associated with higher steady state transcript abundance
of this gene (Figure \ref{fig5}D, E). Transcription appeared to begin at the TE
insertion point, rather than the transcriptional start site of the gene
(Figure \ref{fig5}D). Accessions containing the TE insertion had significantly
higher AT2G01360 transcript abundance than the accessions without the TE
insertion (p \textless{} $1.8\times10^{-7}$, Mann-Whitney U test). The apparent
transcriptional activation, linked with presence of a TE belonging to
the \emph{HELITRON1} family, indicates that this element may carry
sequences or other regulatory information that has altered the
expression of genes downstream of the TE insertion site. Importantly,
this variant was classified as a low-LD TE insertion, as it is not in LD
with surrounding SNPs, and therefore the associated changes in gene
transcript abundance would not be identified using only SNP data. This
TE variant was also upstream of \emph{QPT }(AT2G01350), involved in NAD
biosynthesis \cite{Katoh:2006if}, which did not show alterations in
steady state transcript abundance associated with the presence of the TE
insertion, indicating a potential directionality of regulatory elements
carried by the TE (Figure \ref{fig5}D, E). Overall, these examples demonstrate
that TE variants can have unpredictable, yet important, effects on the
expression of nearby genes, and these effects may be missed by studies
focused on genetic variation at the level of SNPs.

\subsection{TE variants explain many DNA methylation differences between
accessions}

As TEs are frequently highly methylated in Arabidopsis
\cite{Lister:2008bh, Cokus:2008fc, Zhang:2006eo, Zilberman:2006hx}, we
next assessed the DNA methylation state surrounding TE variant sites to
determine whether TE variants might be responsible for some of the
differences in DNA methylation patterns previously observed between the
wild accessions \cite{Schmitz:2013iu}. We found that TE variants were
often physically close to DMRs (Figure \ref{fig6}A). Furthermore, C-DMRs were
more often close to a TE insertion than expected, while they were rarely
near a TE deletion (Table \ref{table3}). CG-DMRs were rarely close to TE insertions
or TE deletions. Overall, we found 48\% of the 13,482 previously
reported population C-DMRs were located within 1 kb of a TE variant
(predominantly TE insertions), while only 15\% of CG-DMRs were within 1
kb of a TE variant (Table \ref{table3}). For C-DMRs, this was significantly more
than expected by chance, while it was significantly less than expected
for CG-DMRs (p \textless{} $1\times10^{-4}$, determined by resampling 10,000
times). To determine if DMR methylation levels were dependent on the
presence/absence of nearby TE variants, we calculated Pearson
correlation coefficients between the DNA methylation level at each DMR
and the presence/absence of the nearest TE variant. We observed a
negative correlation between the distance from a C-DMR to the nearest TE
insertion and the correlation between the DNA methylation level at the
C-DMR with the presence/absence of the TE insertion (Figure \ref{fig6}B). This
suggests a distance-dependent effect of TE insertion presence on C-DMR
methylation. In contrast, we found no such relationship for TE deletions
on C-DMRs, or for insertions or deletions on CG-DMRs (Figure \ref{fig6}B). DNA
methylation levels at C-DMRs located within 1 kb of a TE insertion
(TE-DMRs) were more often positively correlated with the
presence/absence of a TE insertion than the DNA methylation levels at
C-DMRs further than 1 kb from a TE insertion (non-TE-DMRs). This was
evident from the distribution of correlations for non-TE-DMRs being
centred around zero, whereas for TE-DMRs this distribution was skewed to
the right (Figure \ref{fig6}C, \emph{D}=0.23). For TE deletions, we did not
observe such a difference in the distributions of correlation
coefficients between TE-DMRs and non-TE-DMRs, nor for CG-DMRs and their
nearby TE insertions or deletions (Figure \ref{fig6}C, \emph{D=}0.07-0.10).
Furthermore, DNA methylation levels were often higher in the presence of
the nearby TE insertion, while this relationship was generally not
observed for C-DMRs further than 1 kb from a TE variant, for TE
deletions, or for CG-DMRs (Figure 6 - figure supplement \hyperref[fig6s1]{1}).

As the above correlations between TE presence/absence and DMR
methylation level rely on the TE variants having a high MAF, this
precludes an analysis of the effect of rare variants on DMR methylation
levels. To determine the effect that these rare TE variants may have on
DMR methylation levels, we performed a burden test for enrichment of DMR
methylation extremes at TE-DMRs, similar to as was done to test the
effect of rare variants on gene expression. We found a strong enrichment
for high C-DMR and CG-DMR methylation level ranks for TE insertions,
while TE deletions were associated with both high and low extremes of
DNA methylation levels at C-DMRs, and less so at CG-DMRs (Figure \ref{fig6}D).
This further indicates that the presence of a TE insertion is associated
with higher C-DMR methylation levels, while TE deletions seem to have
more variable effects on DMR methylation levels. This enrichment was
completely absent after repeating the analysis with randomized accession
names (Figure 6 - figure supplement \hyperref[fig6s2]{2}). We also observed a slight
enrichment for low DMR methylation ranks for TE insertions near CG-DMRs,
indicating that the insertion of a TE was sometimes associated with
reduced CG methylation in nearby regions (\textless{}1 kb from the TE).
We examined these TE insertions in a genome browser, and found that some
TE insertions were associated with decreased transcript abundance of
nearby genes, with a corresponding loss of gene body methylation,
offering a potential explanation for the decreased CG methylation
observed near some TE insertions (Figure 6 - figure supplement \hyperref[fig6s3]{3}).

We next examined levels of DNA methylation in regions flanking all TE
variants regardless of the presence or absence of a population DMR call.
While DNA methylation levels around pericentromeric TE insertions and
deletions (\textless{}3 Mb from a centromere) seemed to be unaffected by
the presence of a TE insertion (Figure \ref{fig7}A), TE insertions in the
chromosome arms were associated with an increase in DNA methylation
levels in all contexts (Figure \ref{fig7}A, B). In contrast, TE deletions in the
chromosome arms did not affect patterns of DNA methylation, as the
flanking methylation level in all contexts seemed to remain high
following deletion of the TE (Figure \ref{fig7}A, C). As the change in DNA
methylation levels around TE variant sites appeared to be restricted to
regions \textless{}200 bp from the insertion site, we correlated DNA
methylation levels in 200 bp regions flanking TE variants with the
presence/absence of TE the variants. DNA methylation levels were often
positively correlated with the presence of a TE insertion when the
insertion was distant from a centromere (Figure \ref{fig7}D). TE deletions were
more variably correlated with local DNA methylation levels, but also
showed a slight bias towards positive correlations for TE deletions
distant from the centromeres. As methylome data was available for both
leaf and bud tissue for 12 accessions, we repeated this analysis
comparing between tissue types, but did not observe any difference in
the patterns of methylation surrounding TE variant sites between the two
tissues (Figure 7 - figure supplement \hyperref[fig7s1]{1}).

These results indicate that local DNA methylation patterns are
influenced by the differential TE content between genomes, and that the
DNA methylation-dependent silencing of TEs may lead to formation of DMRs
between wild Arabidopsis accessions. TE insertions appear to be
important in defining local patterns of DNA methylation, while DNA
methylation levels often remain elevated following a TE deletion, and so
are independent from the presence or absence of TEs in these cases.
Importantly, the distance from a TE insertion to the centromere appears
to have a strong impact on whether an alteration of local DNA
methylation patterns will occur. This is likely due to flanking
sequences being highly methylated in the pericentromeric regions, and so
the insertion of a TE cannot further increase levels of DNA methylation.
Overall, a large fraction of the population C-DMRs previously identified
between wild accessions are correlated with the presence of local TE
insertions, but not TE deletions. CG-DMR methylation levels seem to be
mostly independent from the presence/absence of common TEs variants,
while rare TE variants have an impact on DNA methylation levels at both
C-DMRs and CG-DMRs.

\subsection{Genome-wide analysis of TE variant association with methylation
levels highlights distant and local control}

To better quantify the above results, an association scan was conducted
for all common TE variants (\textgreater{}3\% MAF) and all population
C-DMRs for the 124 accessions with both DNA methylation and TE variant
data available. To test the significance of each pairwise correlation,
we collected bootstrap p-value estimates based on 500 permutations of
accession labels. TE-DMR associations were deemed significant if they
had a real association more extreme than any of the 500 permutations (p
\textless{} 1/500). A band of significant associations was observed for
TE insertions and their nearby C-DMRs, signifying a local association
between TE insertion presence/absence and C-DMR methylation (Figure \ref{fig8}A).
This local association was not as strong for TE deletions (Figure \ref{fig8}B),
consistent with our above findings. While TE variants and DNA
methylation showed a local association, it is also possible that TE
variation may influence DNA methylation state more broadly in the
genome, perhaps through production of \emph{trans-}acting smRNAs or
inactivation of genes involved in DNA methylation establishment or
maintenance. To identify any potential enrichment of C-DMRs regulated in
\emph{trans}, we summed the total number of significant associations for
each TE variant across the whole genome (Figure \ref{fig8}A and B, top panels).
At many sites, we found far more significant associations than expected
due to the false positive rate alone. This suggested the existence of
many putative \emph{trans }associations between TE variants and
genome-wide C-DMR methylation levels. We further examined these C-DMRs
that appeared to be associated with a TE insertion in \emph{trans},
checking for TE insertions near these C-DMRs that were present in the
same accessions as the \emph{trans }associated TE, as these could lead
to a false \emph{trans }association. These were extremely rare, with
only 4 such cases for TE insertions, and 38 cases for TE deletions, and
so were unable to explain the high degree of \emph{trans }associations
found.

\section{Discussion}

Here we discovered widespread differential TE content between wild
Arabidopsis accessions, and explore the impact of these variants at the
level of individual accessions. Most TE variants were due to the
\emph{de novo }insertion of TEs, while a smaller subset was likely due
to the deletion of ancestral TE copies, mostly around the
pericentromeric regions. A subset (32\%) of TE variants with a minor
allele frequency above 3\% were able to be tested for linkage with
nearby SNPs. The majority of these TE variants were not in LD with SNPs,
indicating that they represent genetic variants currently overlooked in
genomic studies. We found a marked depletion of TE variants within gene
bodies and DNase I hypersensitivity sites (putative regulatory regions),
indicating that the more deleterious TE insertions may have been removed
from this population through selection. Of those TE variants found in
gene bodies, TE deletions were overrepresented, indicating that the loss
of ancestral TEs inserted within genes may be more frequent, or perhaps
less deleterious, than the \emph{de novo }insertion of TEs into genes.

The identification of a large number of TE variants in this population
gave an opportunity to form statistically robust correlations between TE
presence/absence and transcript abundance from nearby genes, as well as
genome-wide patterns of DNA methylation. We were able to identify
examples where TE variants appear to have an effect upon gene
expression, both in the disruption of transcription and in the spreading
or disruption of regulatory information leading to the transcriptional
activation of genes, indicating that these TE variants can have
important consequences upon the expression of protein coding genes
(Figure \ref{fig5}). In one case, these changes in gene expression could be
linked with phenotypic changes, with accessions containing a TE
insertion more frequently sensitive to bacterial infection. Further
experiments will be needed to establish a causal link between this TE
insertion and the associated phenotype. An analysis of rare TE variants,
present at a low MAF, further strengthened this relationship between TE
presence/absence and altered transcript abundance, as we were able to
identify a strong enrichment of rare TE variants in accessions with
extreme gene expression ranks in the population.

Perhaps most importantly, we provide evidence that differential TE
content between genomes of Arabidopsis accessions underlies a large
fraction of the previously reported population C-DMRs, in agreement with
recent similar findings \cite{Quadrana:2016bi}. Thus, the frequency of
pure epialleles, independent of underlying genetic variation, may be
even more rare than previously anticipated \cite{Richards:2006br}. We
did not find evidence of CG-DMR methylation level being altered by the
presence of common TE variants, but rather rare TE variants may be more
important in shaping patterns of DNA methylation at some CG-DMRs, though
the reasons for this distinction remain unclear. The level of local DNA
methylation changes associated with TE variants was also related to the
distance from a TE variant to the centromere, with variants in the
chromosome arms being more strongly correlated with DNA methylation
levels. This seems to be due to a higher baseline level of DNA
methylation at the pericentromeric regions, which prevent any further
increase in DNA methylation level following insertion of a TE.
Furthermore, we found an important distinction between TE insertions and
TE deletions in the effect that these variants have on nearby DNA
methylation levels. While flanking DNA methylation levels appeared to
increase following a TE insertion, the deletion of an ancestral TE was
often not associated with a corresponding decrease in flanking DNA
methylation levels (Figure \ref{fig7}). This indicates that high levels of DNA
methylation, once established, may be maintained in the absence of the
TE insertion that presumably triggered the original change in DNA
methylation level. It is then possible that TE variants explain more of
the variation in DNA methylation patterns than we find direct evidence
for, if some C-DMRs were formed by the insertion of an ancestral TE that
is now absent in all the accessions analysed here. These DMRs would then
represent the epigenetic ``scars'' of past TE insertions.

Finally, we performed a genome-wide scan of common TE variant
association with C-DMR methylation levels, and found further evidence of
a strong local association between TE insertion presence/absence and
C-DMR methylation level (Figure \ref{fig8}). We were also able to identify some
TE variants that appeared to be associated with changes in DNA
methylation levels at multiple loci throughout the genome, indicating a
possible \emph{trans} regulation of DNA methylation levels linked to
certain TE variants. Further experiments will be required to confirm and
examine the role of these TE variants in determining genome-wide
patterns of DNA methylation. Overall our results show that TE
presence/absence variants between wild Arabidopsis accessions not only
have important effects on nearby gene expression, but can also have a
role in determining local patterns of DNA methylation, and explain many
regions of differential DNA methylation previously observed in the
population.

\section{Methods}

\subsection{TEPID development}

\emph{Mapping}

FASTQ files are mapped to the reference genome using the `tepid-map'
algorithm (Figure \ref{fig1}). This first calls bowtie2 \cite{Langmead:2012jh}
with the following options: `--local', `--dovetail', `--fr', `-R5',
`-N1'. Soft-clipped and unmapped reads are extracted using Samblaster
\cite{Faust:2014hf}, and remapped using the split read mapper Yaha
\cite{Faust:2012ic}, with the following options: `-L 11', `-H 2000', `-M
15', `-osh'. Split reads are extracted from the Yaha alignment using
Samblaster \cite{Faust:2014hf}. Alignments are then converted to bam
format, sorted, and indexed using samtools \cite{Li:2009ka}.

\emph{TE variant discovery}

The `tepid-discover' algorithm examines mapped bam files generated by
the `tepid-map' step to identify TE presence/absence variants with
respect to the reference genome. Firstly, mean sequencing coverage, mean
library insert size, and standard deviation of the library insert size
is estimated. Discordant read pairs are then extracted, defined as mate
pairs that map more than 4 standard deviations from the mean insert size
from one another, or on separate chromosomes.

To identify TE insertions with respect to the reference genome, split
read alignments are first filtered to remove reads where the distance
between split mapping loci is less than 5 kb, to remove split reads due
to small indels, or split reads with a mapping quality (MAPQ) less than
5. Split and discordant read mapping coordinates are then intersected
using pybedtools \cite{Dale:2011cl, Quinlan:2010km} with the Col-0
reference TE annotation, requiring 80\% overlap between TE and read
mapping coordinates. To determine putative TE insertion sites, regions
are then identified that contain independent discordant read pairs
aligned in an orientation facing one another at the insertion site, with
their mate pairs intersecting with the same TE (Figure \ref{fig1}). The total
number of split and discordant reads intersecting the insertion site and
the TE is then calculated, and a TE insertion predicted where the
combined number of reads is greater than a threshold determined by the
average sequencing depth over the whole genome (1/10 coverage if
coverage is greater than 10, otherwise a minimum of 2 reads).
Alternatively, in the absence of discordant reads mapped in orientations
facing one another, the required total number of split and discordant
reads at the insertion site linked to the inserted TE is set higher,
requiring twice as many reads.

To identify TE absence variants with respect to the reference genome,
split and discordant reads separated \textgreater{}20 kb from one
another are first removed, as 99.9\% of Arabidopsis TEs are shorter than
20 kb, and this removes split reads due to larger structural variants
not related to TE diversity. Col-0 reference annotation TEs that
are located within the genomic region
spanned by the split and discordant reads are then identified. TE
absence variants are predicted where at least 80\% of the TE sequence is
spanned by a split or discordant read, and the sequencing depth within
the spanned region is \textless{}10\% the sequencing depth of the 2 kb
flanking sequence, and there are a minimum number of split and
discordant reads present, determined by the sequencing depth (1/10
coverage; Figure \ref{fig1}). A threshold of 80\% TE sequence spanned by split
or discordant reads is used, as opposed to 100\%, to account for
misannotation of TE sequence boundaries in the Col-0 reference TE
annotation, as well as TE fragments left behind by DNA TEs during
cut-paste transposition (TE footprints) that may affect the mapping of
reads around annotated TE borders \cite{Plasterk:1991vd}. This was found
to improve TE absence detection using simulated data. Furthermore, the
coverage within the spanned region may be more than 10\% that of the
flanking sequence, but in such cases twice as many split and discordant
reads are required. If multiple TEs are spanned by the split and
discordant reads, and the above requirements are met, multiple TEs in
the same region can be identified as absent with respect to the
reference genome. Absence variants in non-Col-0 accessions are
subsequently recategorized as TE insertions present in the Col-0 genome
but absent from a given wild accession.

\emph{TE variant refinement}

Once TE insertions are identified using the `tepid-map' and
`tepid-discover' algorithms, these variants can be refined if multiple
related samples are analysed. The `tepid-refine' algorithm is designed
to interrogate regions of the genome in which a TE insertion was
discovered in other samples but not the sample in question, and check
for evidence of that TE insertion in the sample using lower read count
thresholds compared to the `tepid-discover' step. In this way, the
refine step leverages TE variant information for a group of related
samples to reduce false negative calls within the group. This
distinguishes TEPID from other similar methods for TE variant discovery
utilizing short sequencing reads. A file containing the coordinates of
each insertion, and a list of sample names containing the TE insertion
must be provided to the `tepid-refine' algorithm, which this can be
generated using the `merge\_insertions.py' script included in the TEPID
package. Each sample is examined in regions where there was a TE
insertion identified in another sample in the group. If there is a
sequencing breakpoint within this region (no continuous read coverage
spanning the region), split reads mapped to this region will be
extracted from the alignment file and their coordinates intersected with
the TE reference annotation. If there are split reads present at the
variant site that are linked to the same TE as was identified as an
insertion at that location, this TE insertion is recorded in a new file
as being present in the sample in question. If there is no sequencing
coverage in the queried region for a sample, an ``NA'' call is made
indicating that it is unknown whether the particular sample contains the
TE insertion or not.

While the above description relates specifically to use of TEPID for
identification of TE variants in Arabidopsis in this study, this method
can be also applied to other species, with the only prerequisite being
the annotation of TEs in a reference genome and the availability of
paired-end DNA sequencing data.

\subsection{TE variant simulation}

To test the sensitivity and specificity of TEPID, 100 TE insertions (50
copy-paste transpositions, 50 cut-paste transpositions) and 100 TE
absence variants were simulated in the Arabidopsis genome using the
RSVSim R package, version 1.7.2 \cite{Bartenhagen:2013kl}, and synthetic
reads generated from the modified genome at various levels of sequencing
coverage using wgsim \cite{Li:2009ka} (\url{https://github.com/lh3/wgsim}).
These reads were then used to calculate the true positive, false
positive, and false negative TE variant discovery rates for TEPID at
various sequencing depths, by running `tepid-map' and `tepid-discover'
using the simulated reads with the default parameters (Figure 1 - figure
supplement \hyperref[fig1s1]{1}).

\subsection{Estimation of sensitivity}

Previously published 100 bp paired end sequencing data for L\emph{er}
(\url{http://1001genomes.org/data/MPI/MPISchneeberger2011/releases/current/Ler-1/Reads/}
; \cite{Schneeberger:2011ft}) was downloaded and analyzed with the TEPID
package to identify TE variants. Reads providing evidence for TE
variants were then mapped to the \emph{de novo} assembled L\emph{er}
genome \cite{Chin:2013iw}. To determine whether reads mapped to
homologous regions of the L\emph{er} and Col-0 reference genome, the
\emph{de novo }assembled L\emph{er }genome sequence between mate pair
mapping locations in L\emph{er} were extracted, with repeats masked
using RepeatMasker with RepBase-derived libraries and the default
parameters (version 4.0.5, \url{http://www.repeatmasker.org}). A blastn
search was then conducted against the Col-0 genome using the following
parameters: `-max-target-seqs 1', `-evalue 1e-6' \cite{Camacho:2009fc}.
Coordinates of the top blast hit for each read location were then
compared with the TE variant sites identified using those reads. To
estimate false negative rates for TEPID TE absence calls, L\emph{er} TE
absence calls were compared with a known set of Col-0-specific TE
insertions, absent in L\emph{er }\cite{Quadrana:2016bi}. For TEPID TE
insertion calls, we mapped Col-0 DNA sequencing reads
\cite{Jiang:2014ih} to the L\emph{er }PacBio assembly, and identified
sites with read evidence reaching the TEPID threshold for a TE insertion
call to be made.

\subsection{Arabidopsis TE variant discovery}

We ran the TEPID, including the insertion refinement step, on previously
published sequencing data for 216 different Arabidopsis populations
(NCBI SRA SRA012474; Schmitz et al. 2013), mapping to the TAIR10
reference genome and using the TAIR9 TE annotation. The `--mask' option
was also used to mask the mitochondrial and plastid genomes. We also ran
TEPID using previously published transgenerational data for salt stress
and control conditions (NCBI SRA SRP045804; \cite{Jiang:2014ih}), using
the `--mask' option to mask mitochondrial and plastid genomes, and the
`--strict' option for highly related samples.

\subsection{TE variant / SNP comparison}

SNP information for 216 Arabidopsis accessions was obtained from the
1001 genomes data center
(\url{http://1001genomes.org/data/Salk/releases/2013_24_01/}; \cite{Schmitz:2013iu}).
This was formatted into reference (Col-0 state), alternate,
or NA calls for each SNP. Accessions with both TE variant information
and SNP data were selected for analysis. Hierarchical clustering of
accessions by SNPs as well as TE variants were used to identify
essentially clonal accessions, as these would skew minor allele
frequency calculations. A single representative from each cluster of
similar accessions was kept, leading to a total of 187 accessions for
comparison. For each TE variant with minor allele frequency greater than
3\%, the nearest 300 upstream and 300 downstream SNPs with a minor
allele frequency greater than 3\% were selected. Pairwise genotype
correlations (r\textsuperscript{2} values) for all complete cases were obtained for
SNP-SNP and SNP-TE variant states. r\textsuperscript{2} values were then ordered by
decreasing rank and a median SNP-SNP rank value was calculated. For each
of the 600 ranked surrounding positions, the number of times the TE rank
was greater than the SNP-SNP median rank was calculated as a relative LD
metric of TE to SNP. TE variants with less than 200 ranks over the
SNP-SNP median were classified as low-LD insertions. TE variants with
ranks between 200 and 400 were classified as mid-LD, while TE variants
with greater than 400 ranks above their respective SNP-SNP median value
were classified as variants in high LD with flanking SNPs.

\subsection{PCR validations}

\emph{Selection of accessions to be genotyped}

To assess the accuracy of TE variant calls in accessions with a range of
sequencing depths of coverage, we grouped accessions into quartiles
based on sequencing depth of coverage and randomly selected a total of
14 accessions for PCR validations from these quartiles. DNA was
extracted for these accessions using Edward's extraction protocol
\cite{Edwards:1991wp}, and purified prior to PCR using AMPure beads.

\emph{Selection of TE variants for validation and primer design}

Ten TE insertion sites and 10 TE absence sites were randomly selected
for validation by PCR amplification. Only insertions and absence
variants that were variable in at least two of the fourteen accessions
selected to be genotyped were considered. For insertion sites, primers
were designed to span the predicted TE insertion site. For TE absence
sites, two primer sets were designed; one primer set to span the TE,
and another primer set with one primer annealing within the TE
sequence predicted to be absent, and the other primer annealing in the
flanking sequence (Figure 2 - figure supplement \hyperref[fig2s3]{3}). Primer
sequences were designed that did not anneal to regions of the genome
containing previously identified SNPs in any of the 216 accessions
\cite{Schmitz:2013iu} or small insertions and deletions, identified
using lumpy-sv with the default settings
\cite{Layer:2014ie}(\url{https://github.com/arq5x/lumpy-sv}), had an
annealing temperature close to 52ºC calculated based on nearest
neighbor thermodynamics (MeltingTemp submodule in the SeqUtils python
module; \cite{Cock:2009hj}), GC content between 40\% and 60\%, and
contained the same base repeated not more than four times in a
row. Primers were aligned to the TAIR10 reference genome using bowtie2
\cite{Langmead:2012jh} with the `-a' flag set to report all
alignments, and those with more than 5 mapping locations in the genome
were then removed.

\emph{PCR}

PCR was performed with 10 ng of extracted, purified Arabidopsis DNA
using Taq polymerase. PCR products were analysed by agarose gel
electrophoresis. Col-0 was used as a positive control, water was added
to reactions as a negative control.

\subsection{mRNA analysis}

Processed mRNA data for 144 wild Arabidopsis accessions were downloaded
from NCBI GEO GSE43858 \cite{Schmitz:2013iu}. To find differential gene
expression dependent on TE presence/absence variation, we first filtered
TE variants to include only those where the TE variant was shared by at
least 5 accessions with RNA data available. We then grouped accessions
based on TE presence/absence variants, and performed a Mann-Whitney U
test to determine differences in RNA transcript abundance levels between
the groups. We used q-value estimation to correct for multiple testing,
using the R qvalue package v2.2.2 with the following parameters: lambda
= seq(0, 0.6, 0.05), smooth.df = 4 \cite{Storey:2003cj}. Genes were
defined as differentially expressed where there was a greater than 2
fold difference in expression between the groups, with a q-value less
than 0.01. Gene ontology enrichment analysis was performed using PANTHER
(\url{http://pantherdb.org}).

\subsection{DNA methylation data analysis}

Processed base-resolution DNA methylation data for wild Arabidopsis
accessions were downloaded from NCBI GEO GSE43857 \cite{Schmitz:2013iu},
and used to construct MySQL tables in a database.

\subsection{Rare variant analysis}

To assess the effect of rare TE variants on gene expression or DMR DNA
methylation levels, we tested for a burden of rare variants in the
population extremes, essentially as described previously
\cite{Zhao:2016gc}. For each rare TE variant near a gene or DMR, we
ranked the gene expression level or DMR DNA methylation level for all
accessions in the population, and tallied the ranks of accessions
containing a rare variant. These rank counts were then binned to produce
a histogram of the distribution of ranks. We then fit a quadratic model
to the counts data, and calculated the R\textsuperscript{2} and p-value for the fit of
the model.

\subsection{TE variant and DMR genome-wide association analysis}

Accessions were subset to those with both leaf DNA methylation data and
TEPID calls. Pairwise correlations were performed for observed data
pairs for each TE variant and a filtered set of population C-DMRs, with
those C-DMRs removed where more than 15\% of the accessions had no
coverage. This amounted to a final set of 9,777 C-DMRs. Accession names
were then permuted to produce a randomized dataset, and pairwise
correlations again calculated. This was repeated 500 times to produce a
distribution of expected Pearson correlation coefficients for each
pairwise comparison. Correlation values more extreme than any of the 500
permutations were deemed significant.

\section{Data access}

TEPID source code can be accessed at
\url{https://github.com/ListerLab/TEPID}. Code and data needed to
reproduce this analysis can be found at
\url{https://github.com/timoast/Arabidopsis-TE-variants}. L\emph{er }TE
variants are available in Figure 1 - source data 1 and 2. TE variants
identified among the 216 wild Arabidopsis accessions resequenced by
Schmitz et al. (2013) are available in Figure 2 - source data 1, 2 and
3.

\section{Acknowledgments}

This work was supported by the Australian Research Council (ARC) Centre
of Excellence program in Plant Energy Biology CE140100008 (J.B., R.L.).
R.L. was supported by an ARC Future Fellowship (FT120100862) and Sylvia
and Charles Viertel Senior Medical Research Fellowship, and work in the
laboratory of R.L. was funded by the Australian Research Council. T.S.
was supported by the Jean Rogerson Postgraduate Scholarship. S.R.E. was
supported by an Australian Research Council Discovery Early Career
Research Award (DE150101206). We thank Robert J. Schmitz, Mathew G.
Lewsey, Ronan C. O'Malley, and Ian Small for their critical reading of
the manuscript, and Kevin Murray for his helpful comments regarding the
development of TEPID.

\section{Author contributions}

R.L. and T.S. designed the research project. R.L. and J.B. supervised
research. T.S. developed and tested TEPID. J.C. performed PCR
validations of TE variants. T.S. and S.R.E. performed bioinformatic
analysis. Y.K. provided statistical guidance. R.L., T.S., J.B. and
S.R.E. prepared the manuscript.

\section{Competing financial interests}

The authors declare no competing financial interests.

\pagebreak

\printbibliography

%\bibliography{references}{}
%\bibliographystyle{ieeetr}

% Figure 1

\pagebreak

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\paperwidth]{Figure1/Figure1.png}
  \caption{TE variant discovery pipeline}
  \label{fig1}
\end{figure}

Principle of TE variant discovery using split and discordant read
mapping positions. Paired end reads are first mapped to the reference
genome using Bowtie2 \cite{Langmead:2012jh}. Soft-clipped or unmapped
reads are then extracted from the alignment and re-mapped using Yaha, a
split read mapper \cite{Faust:2012ic}. All read alignments are then used
by TEPID to discover TE variants relative to the reference genome, in
the `tepid-discover' step. When analyzing groups of related samples,
these variants can be further refined using the `tepid-refine' step,
which examines in more detail the genomic regions where there was a TE
variant identified in another sample, and calls the same variant for the
sample in question using lower read count thresholds as compared to the
`tepid-discover' step, in order to reduce false negative variant calls
within a group of related samples.

\pagebreak

% Figure 1 - figure supplement 1

%\setcounter{figure}[0]

\begin{suppfigure}[h]
  \centering
  \includegraphics[width=0.7\paperwidth]{Figure1/Figure1-figure_supplement1.png}
  \caption{figure supplement 1}
  \label{fig1s1}
\end{suppfigure}

Testing of the TEPID pipeline using simulated TE variants in the
Arabidopsis Col-0 genome (TAIR10), for a range of sequencing coverage
levels. TE insertions (A) and TE absence calls (B).

\pagebreak

% Figure 2

%\setcounter{figure}[2]

\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth]{Figure2/Figure2.png}
  \caption{Extensive novel genetic diversity uncovered by TE variant analysis}
  \label{fig2}
\end{figure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Distribution of identified TE variants on chromosome 1, with
  distributions of all Col-0 genes, Col-0 TEs, and population DMRs.
\item
  Frequency of TE variants at different genomic features.
\item
  Proportion of TE variants within each genomic feature classified as
  deletions or insertions.
\item
  Proportion of TE variants within each genomic feature classified as
  rare or common.
\item
  Enrichment and depletion of TE variants categorized by TE superfamily
  compared to the expected frequency due to genomic occurrence.
\end{enumerate}

\pagebreak

% Figure 2 - figure supplement 1

\setcounter{suppfigure}{1}
\begin{suppfigure}
  \centering
  \includegraphics{Figure2/Figure2-figure_supplement_1.png}
  \caption{figure supplement 1}
  \label{fig2s1}
\end{suppfigure}

Number of additional TE insertion calls made due to the TEPID refinement
step for each accession in the population.

\pagebreak
% Figure 2 - figure supplement 2
\setcounter{suppfigure}{1}

\begin{suppfigure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure2/Figure2-figure_supplement_2.png}
  \caption{figure supplement 2}
  \label{fig2s2}
\end{suppfigure}

Minor allele frequency distribution for TE variants.

\pagebreak
% Figure 2 - figure supplement 3
\setcounter{suppfigure}{1}

\begin{suppfigure}
  \centering
  \includegraphics[width=1\textwidth]{Figure2/Figure2-figure_supplement_3.png}
  \caption{figure supplement 3}
  \label{fig2s3}
\end{suppfigure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  PCR validations for a TE absence variant. Accessions that were
  predicted to contain a TE insertion or TE absence are marked in bold.
  Two primer sets were used; forward (F) and reverse (R) or internal
  (I). Accessions with a TE absence will not produce the FI band and
  produce a shorter FR product, with the change in size matching the
  size of the deleted TE.
\item
  PCR validations for a TE insertion variant. One primer set was used,
  spanning the TE insertion site. A band shift of approximately 200 bp
  can be seen, corresponding to the size of the inserted TE.
\end{enumerate}

\pagebreak
% Figure 2 - figure supplement 4

\setcounter{suppfigure}{1}

\begin{suppfigure}
  \centering
  \includegraphics[width=1\textwidth]{Figure2/Figure2-figure_supplement_4.pdf}
  \caption{figure supplement 4}
  \label{fig2s4}
\end{suppfigure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Number of TE absence variants identified versus the sequencing depth
  of coverage for each accession.
\item
  Number of TE insertion variants identified versus the sequencing depth
  of coverage for each accession.
\item
  Number of additional TE insertion calls made due to the TEPID
  refinement step versus sequencing depth of coverage for all
  accessions.
\end{enumerate}

\pagebreak
% Figure 2 - figure supplement 5

\setcounter{suppfigure}{1}

\begin{suppfigure}
  \centering
  \includegraphics[width=0.7\textwidth]{Figure2/Figure2-figure_supplement_5.png}
  \caption{figure supplement 5. Frequency of TE insertion in the \emph{KNOT} region}
  \label{fig2s5}
\end{suppfigure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Number of TE insertion variants within each 300 kb \emph{KNOT ENGAGED ELEMENT (KEE)},
  vertical lines) and the number of TE insertion variants
  found in 10,000 randomly selected 300 kb windows (histogram).
\item
  Table showing number of TE insertion variants within each \emph{KEE}
  region, and the associated p-value determined by resampling 10,000
  times.
\end{enumerate}

\pagebreak

% Figure 2 - figure supplement 6
\setcounter{suppfigure}{1}

\begin{suppfigure}
  \centering
  \includegraphics[width=0.7\textwidth]{Figure2/Figure2-figure_supplement_6.png}
  \caption{figure supplement 6. Length distribution for all Col-0 TEs and all TE variants}
  \label{fig2s6}
\end{suppfigure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Histogram showing lengths of all annotated TEs in the Col-0 reference
  genome.
\item
  Histogram showing lengths of all TE variants.
\item
  Density distribution of log10 TE length for all Col-0 TEs (red) and TE
  variants (blue).
\end{enumerate}

\pagebreak

% Figure 2 - figure supplement 7
\setcounter{suppfigure}{1}

\begin{suppfigure}[!ht]
  \centering
  \includegraphics[width=1\textwidth]{Figure2/Figure2-figure_supplement_7.png}
  \caption{figure supplement 7}
  \label{fig2s7}
\end{suppfigure}

TE family enrichments and depletions for TE insertions and TE deletions.

\pagebreak

% Figure 3

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.5\textwidth]{Figure3/Figure3.png}
  \caption{Patterns of TE-SNP linkage}
  \label{fig3}
\end{figure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  r\textsuperscript{2} correlation matrices for individual representative high and low-LD
  TE variants showing the background level of SNP-SNP linkage.
\item
  Rank order plots for individual representative high and low-LD TE
  variants (matching those shown in A). Red line indicates the median r\textsuperscript{2}
  value for each rank across SNP-based values. Blue line indicates r\textsuperscript{2}
  values for TE-SNP comparisons. Grey lines indicate all individual
  SNP-SNP comparisons.
\item
  Histogram of the number of TE r\textsuperscript{2} ranks (0-600) that are above the
  SNP-based median r\textsuperscript{2} value for testable TE variants.
\item
  Boxplots showing distribution of minor allele frequencies for each LD
  category. Boxes represent the interquartile range (IQR) from quartile
  1 to quartile 3. Boxplot upper whiskers represent the maximum value,
  or the upper value of the quartile 3 plus 1.5 times the IQR (whichever
  is smaller). Boxplot lower whisker represents the minimum value, or
  the lower value of the quartile 1 minus 1.5 times the IQR (whichever
  is larger).
\end{enumerate}

(E). Proportion of TE insertions, TE deletions, and unclassified TE
variants in each LD category.

\pagebreak

% Figure 3 - figure supplement 1

\setcounter{suppfigure}{2}

\begin{suppfigure}
  \centering
  \includegraphics[width=0.7\textwidth]{Figure3/Figure3-figure_supplement1.png}
  \caption{figure supplement 1}
  \label{fig3s1}
\end{suppfigure}

Distribution of TE variants across chromosome 1 for each LD category
(high, mid, low).

\pagebreak

% Figure 4

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{Figure4/Figure4.png}
  \caption{Differential transcript abundance associated with TE variant presence/absence}
  \label{fig4}
\end{figure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Volcano plots showing transcript abundance differences for genes
  associated with TE insertion variants at different positions,
  indicated in the plot titles. Genes with significantly different
  transcript abundance in accessions with a TE insertion compared to
  accessions without a TE insertion are colored blue (lower transcript
  abundance in accessions containing TE insertion) or red (higher
  transcript abundance in accessions containing TE insertion). Vertical
  lines indicate ±2 fold change in FPKM. Horizontal line indicates the
  1\% FDR.
\item
  Relationship between TE rare variant counts and gene expression rank.
  Plot shows the cumulative number of rare TE variants in equal-sized
  bins for gene expression ranks, from the lowest-ranked accession
  (left) to the highest-ranked accession (right). Lines indicate the fit
  of a quadratic model.
\end{enumerate}

\pagebreak

% Figure 4 - figure supplement 1

\setcounter{suppfigure}{3}

\begin{suppfigure}
  \centering
  \includegraphics[width=1\textwidth]{Figure4/Figure4-figure_supplement1.png}
  \caption{figure supplement 1}
  \label{fig4s1}
\end{suppfigure}

Relationship between rare TE variants and gene expression rank as for
Figure 4B, for permuted TE variants.

\pagebreak

% Figure 4 - figure supplement 2

\setcounter{suppfigure}{3}

\begin{suppfigure}
  \centering
  \includegraphics[width=1\textwidth]{Figure4/Figure4-figure_supplement2.png}
  \caption{figure supplement 2}
  \label{fig4s2}
\end{suppfigure}

Relationship between rare TE variants and gene expression rank as for
Figure 4B, for TE insertions (A) and TE deletions (B) separately.

\pagebreak

% Figure 5

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\textwidth]{Figure5/Figure5.png}
  \caption{Effects of TE variants on local gene expression}
  \label{fig5}
\end{figure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Genome browser representation of RNA-seq data for genes \emph{AtRLP18}
  (AT2G15040) and a leucine-rich repeat family protein (AT2G15042) for
  Db-1, containing a TE insertion within the exon of the gene
  \emph{AtRLP18}, and for a Col-0 (not containing the TE insertion
  within the exon of \emph{AtRLP18}). Inset shows magnified view of the
  TE insertion site.
\item
  Heatmap showing \emph{AtRLP18 }and AT2G15042 RNA-seq FPKM values for
  all accessions.
\item
  Percentage of accessions with resistance to \emph{Pseudomonas syringae
  }transformed with different \emph{avr }genes, for accessions
  containing or not containing a TE insertion in* AtRLP18.*
\item
  Genome browser representation of RNA-seq data for a PPR
  protein-encoding gene (AT2G01360) and \emph{QPT }(AT2G01350), showing
  transcript abundance for these genes in accessions containing a TE
  insertion variant in the upstream region of these genes.
\item
  Heatmap representation of RNA-seq FPKM values for \emph{QPT }and a
  gene encoding a PPR protein (AT2G01360), for all accessions. Note that
  scales are different for the two heatmaps, due to the higher
  transcript abundance of \emph{QPT} compared to AT2G01360. Scale
  maximum for AT2G01350 is $3.1\times10^5$, and for AT2G01360 is $5.9\times10^4$.
\end{enumerate}

\pagebreak

% Figure 6

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\textwidth]{Figure6/Figure6.png}
  \caption{TE variants are associated with nearby DMR methylation levels}
  \label{fig6}
\end{figure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Distribution of distances from TE variants to the nearest population
  DMR, for TE deletions and TE insertion, C-DMRs and CG-DMRs.
\item
  Pearson correlation between DMR DNA methylation level and TE
  presence/absence, for all DMRs and their closest TE variant, versus
  the distance from the DMR to the TE variant (log scale). Blue lines
  show a linear regression between the correlation coefficients and the
  log10 distance to the TE variant.
\item
  Empirical cumulative distribution of Pearson correlation coefficients
  between TE presence/absence and DMR methylation level for TE
  insertions, TE deletions, C-DMRs and CG-DMRs. The Kolmogorov--Smirnov
  statistic is shown in each plot, indicated by \emph{D.}
\item
  Relationship between rare TE variant counts and nearby DMR DNA
  methylation level ranks, for TE insertions, deletions, C-DMRs, and
  CG-DMRs. Plot shows the cumulative number of rare TE variants in
  equal-sized bins of DMR methylation level ranks, from the lowest
  ranked accession (left) to the highest ranked accession (right). Lines
  indicate the fit of a quadratic model, and the corresponding R\textsuperscript{2} and p
  values are shown in each plot.
\end{enumerate}

\pagebreak

% Figure 6 - figure supplement 1

\setcounter{suppfigure}{5}

\begin{suppfigure}
  \centering
  \includegraphics[width=0.6\textwidth]{Figure6/Figure6-figure_supplement1.png}
  \caption{figure supplement 1}
  \label{fig6s1}
\end{suppfigure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  DNA methylation density distribution at C-DMRs within 1 kb of a TE
  variant (TE-DMRs) or further than 1 kb from a TE variant
  (non-TE-DMRs), in the presence or absence of the TE, for TE insertions
  and TE deletions.
\item
  As for A, for CG-DMRs.
\end{enumerate}

\pagebreak

% Figure 6 - figure supplement 2

\setcounter{suppfigure}{5}

\begin{suppfigure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure6/Figure6-figure_supplement2.png}
  \caption{figure supplement 2}
  \label{fig6s2}
\end{suppfigure}

Cumulative number DMR methylation level ranks for DMRs near rare TE
variants with accessions selected at random. Lines indicate the fit of a
quadratic model, and the corresponding R\textsuperscript{2} and p values are shown in each
plot.

\pagebreak

% Figure 6 - figure supplement 3

\setcounter{suppfigure}{5}

\begin{suppfigure}
  \centering
  \includegraphics[width=1\textwidth]{Figure6/Figure6-figure_supplement3.png}
  \caption{figure supplement 3}
  \label{fig6s3}
\end{suppfigure}

Selected examples of TE insertions apparently associated with
transcriptional downregulation of nearby genes and loss of gene body CG
methylation leading to the formation of a CG-DMR.

\pagebreak

% Figure 7

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{Figure7/Figure7.png}
  \caption{Local patterns of DNA methylation surrounding TE variant sites}
  \label{fig7}
\end{figure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Heatmap showing DNA methylation levels in 200 bp bins flanking TE
  variant sites, +/- 2 kb from the TE insertion point. TE variants were
  grouped into pericentromeric variants (\textless{}3 Mb from a
  centromere) or variants in the chromosome arms (\textgreater{}3 Mb
  from a centromere).
\item
  Line plot showing the DNA methylation level in each sequence context
  for TE insertion sites, +/- 2 kb from the TE insertion point.
\item
  As for B, for TE deletions.
\item
  Distribution of Pearson correlation coefficients between TE
  presence/absence and DNA methylation levels in the 200 bp regions
  flanking TE variant, ordered by distance to the centromere.
\end{enumerate}

\pagebreak

% Figure 7 - figure supplement 1

\setcounter{suppfigure}{6}

\begin{suppfigure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure7/Figure7-figure_supplement1.png}
  \caption{figure supplement 1}
  \label{fig7s1}
\end{suppfigure}

Heatmap showing DNA methylation levels in 200 bp bins flanking TE
variant sites in the 12 accessions with DNA methylation data for both
leaf and bud tissue, +/- 2 kb from the TE insertion point. TE variants
were grouped into pericentromeric variants (\textless{}3 Mb from a
centromere) or variants in the chromosome arms (\textgreater{}3 Mb from
a centromere).

\pagebreak

% Figure 8

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.5\textwidth]{Figure8/Figure8.png}
  \caption{Association scan between TE variants can C-DMR methylation variation}
  \label{fig8}
\end{figure}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  Significant correlations between TE insertions and C-DMR DNA
  methylation level. Points show correlations between individual TE-DMR
  pairs that were more extreme than any of 500 permutations of the DMR
  data. Top plots show the total number of significant correlations for
  each TE insertion across the whole genome.
\item
  As for (A), for TE deletions.
\end{enumerate}

\pagebreak

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Mapping of paired-end reads providing evidence for TE presence/absence
      variants in the L\emph{er} reference genome}
    \label{table1}
    \begin{tabular}{@{}cccccc@{}}
      \toprule
      \textbf{}  & \textbf{Concordant} & \textbf{Discordant} & \textbf{Split} & \textbf{Unmapped} & \textbf{Total} \\ \midrule
      Col-0 mapped & 0 & 993 & 9513 & 0 & 10206 \\
      L\emph{er} mapped & 10073 & 92 & 34 & 7 & 10206 \\ \bottomrule
    \end{tabular}
    \begin{tablenotes}
      \small
    \item Note: Discordant and split read categories are not mutually exclusive, as some discordant reads may have one read in the mate pair split-mapped.
    \end{tablenotes}
  \end{threeparttable}
\end{table}

\vspace*{1 cm}

\begin{table}[h]
  \centering
  \caption{Summary of TE variant classifications}
  \label{table2}
  \begin{tabular}{@{}ccc@{}}
    \toprule
    \textbf{TEPID call}        & \textbf{TE classification} & \textbf{Count} \\ \midrule
    \multirow{3}{*}{Insertion} & NA                         & 310            \\
    & Insertion                  & 14689          \\
    & Deletion                  & 8              \\ \midrule
    \multirow{3}{*}{Absence}   & NA                         & 1852           \\
    & Insertion                  & 388            \\
    & Deletion                   & 5848  \\ \bottomrule
  \end{tabular}
\end{table}

\vspace*{1 cm}

\begin{table}[h]
  \centering
  \caption{Percentage of DMRs within 1 kb of a TE variant}
  \label{table3}
  \begin{tabular}{@{}cccc|ccc@{}}
    \toprule
    \textbf{}              & \multicolumn{3}{c|}{\textbf{C-DMRs}}                     & \multicolumn{3}{c}{\textbf{CG-DMRs}}                    \\ \midrule
    \textbf{}              & \textbf{Observed} & \textbf{Expected} & \textbf{95\% CI} & \textbf{Observed} & \textbf{Expected} & \textbf{95\% CI} \\ \midrule
    \textbf{TE deletions}  & 8.7               & 16                & 0.0078           & 4.3               & 16                & 0.0041           \\
    \textbf{TE insertions} & 36                & 26                & 0.0089           & 9.4               & 26                & 0.0047           \\
    \textbf{NA calls}      & 3.4               & 6.2               & 0.0052           & 1.7               & 6.2               & 0.0027           \\
    \textbf{Total}         & 48                & 41                & 0.01             & 15                & 42                & 0.0054           \\ \bottomrule
  \end{tabular}
\end{table}

\end{document}