### 2016-11-14 AnalysisCode/intergenic_regions.Rmd
Add R code for getting intergenic regions

---
title: "Intergenic regions"
author: "Tim Stuart"
date: "11/11/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(BSgenome.Athaliana.TAIR.TAIR9)
library(readr)
library(dplyr)
```

Load all the tracks

```{r}
genes <- import.gff("../RawData/TAIR10_GFF3_genes.gff.gz")
te <- import("../RawData/TAIR9_TE.bed.gz",
             format = "bed",
             extraCols = c(strand = "character",
                           TE = "character",
                           family = "character",
                           superfamily = "character"))
pseudogene <- import("../ProcessedData/GeneFeatures/pseudogene.bed.gz",
                     format = "bed",
                     extraCols = c(dot = "character",
                                   strand = "character",
                                   source = "character",
                                   type = "character",
                                   dot = "character",
                                   id = "character"))
dhs <- import("../RawData/Sullivan_DHS_PE_peaks_control.bed.gz",
              format = "bed",
              extraCols = c(dot = "character",
                            zero = "numeric",
                            second_dot = "character",
                            a = "numeric",
                            b = "numeric"))
centromeres <- import("../RawData/centromere_positions.txt",
                      format = "bed")

# make chromosome names match the genome
genes <- dropSeqlevels(genes, c("chrM", "chrC"))
genes <- renameSeqlevels(genes, c("Chr1", "Chr2", "Chr3", "Chr4", "Chr5"))
te <- renameSeqlevels(te, c("Chr1", "Chr2", "Chr3", "Chr4", "Chr5"))
pseudogene <- renameSeqlevels(pseudogene, c("Chr1", "Chr2", "Chr3", "Chr4", "Chr5"))
dhs <- renameSeqlevels(dhs, c("Chr1", "Chr2", "Chr3", "Chr4", "Chr5"))
centromeres <- renameSeqlevels(centromeres, c("Chr1", "Chr2", "Chr3", "Chr4", "Chr5"))

# remove metadata columns
mcols(te) <- NULL
mcols(genes) <- NULL
mcols(pseudogene) <- NULL
mcols(dhs) <- NULL
```

Now extend gene coordinates 2 kb in each direction.

```{r extend_coords}
# need to remove entries that cover the entire chromosome, they will overlap everything
chromosome_entries <- width(ranges(genes)) > 10^6

# check we have 5 chromosomes 
sum(chromosome_entries)

# now remove those entries
genes <- genes[!chromosome_entries]

# now extend 2 kb each direction
extended_genes <- genes + 2000
```

Merge all the coordinates and reduce the final genomic ranges object, then take the coordinates between.

```{r}
all_features <- reduce(c(extended_genes, te, pseudogene, dhs, centromeres))
seq_between <- gaps(all_features)
# remove any of length 1
intergenic <- seq_between[!width(seq_between) == 1]
```

Now intersect with the TE variants to find the proportion of all these ranges that contain at least 1 TE variant

First load the TEPID data:

```{r intersect_TEPID}
tepid <- read_tsv("../RawData/TEPID_TEPAV.tsv.gz")
tepid_coords <- makeGRangesFromDataFrame(tepid[,1:3])
tepid_coords <- renameSeqlevels(tepid_coords, c("Chr1", "Chr2", "Chr3", "Chr4", "Chr5"))
```

Find the percentage of all intergenic regions that contain a TE insertion

```{r}
length(subsetByOverlaps(intergenic, tepid_coords)) / length(intergenic) * 100
```

Now try splitting up all intergenic regions longer than 1 kb. If longer than x kb, cut into x+1 equal bins, where x >= 1

```{r split_intergenic}
gr <- GRanges()
for(i in seq_along(intergenic)) {
        l <- width(intergenic[i])
        if(l < 1000) {
                gr <- c(gr, intergenic[i])
        } else {
                gr <- c(gr, unlist(tile(intergenic[i], n = floor(l / 1000))))
        }
}
```

```{r intersect_tiled_intergenic}
length(subsetByOverlaps(gr, tepid_coords)) / length(gr) * 100
```

Let's go with the original, non-cut, data.