### 2016-11-20 AnalysisCode/basho.Rmd
Add basho allele freq comparison

---
title: "basho"
author: "Tim Stuart"
date: "20/11/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basho allele frequencies

Aim: compare the estimated allele frequencies for TEs genotyped by Hollister and Gaut (2007) to my allele frequencies for the same TEs using the 1001 genomes read data.

First, read in the data we need:

```{r libraries, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(GenomicRanges)
library(rtracklayer)
```

```{r load_data}
basho <- read_csv("../RawData/Gaut_basho_supplement.csv", skip = 1)
tepav <- read_tsv("../RawData/TEPID_TEPAV.tsv.gz")
colnames(basho) <- make.names(colnames(basho))
basho_gr <- basho[,c(3:5, 8)] %>%
        makeGRangesFromDataFrame(start.field = "Testart..bp.", end.field = "Teend..bp.",
                                 keep.extra.columns = TRUE)
basho_gr <- renameSeqlevels(basho_gr, c("chr1", "chr2", "chr3", "chr4", "chr5"))

te <- import("../../TEPID/Annotation/Arabidopsis/TAIR9_TE.bed.gz",
             format = "bed",
             extraCols = c(strand = "character",
                           TE = "character",
                           family = "character",
                           superfamily = "character"))
```

For each TE set of coordinates in the basho set, find what TE overlaps and record this information. Then get the number of accessions predicted to contain/lack that TE from my data, and the genotyping results from Gaut.

```{r}
hits <- findOverlaps(te, basho_gr)
basho_hits <- as_data_frame(basho_gr[subjectHits(hits),])
```

```{r}
te_hits <- as_data_frame(te[queryHits(hits),])
```

```{r}
basho_hits$TE <- te_hits$TE
```

Now join with the TEPAV data

```{r}
dat <- left_join(basho_hits, tepav, by = "TE") %>%
        rowwise() %>%
        filter("Col-0" %in% unlist(strsplit(Accessions_TE_present, ","))) %>%
        select(Occ..Freq., Count_TE_present)

p <- ggplot(dat, aes(Occ..Freq., Count_TE_present/216)) + geom_point() +
        geom_smooth(method = "lm", se = FALSE) +
        xlab("Occupation frequency Hollister and Gaut (2007)") +
        ylab("Occupation frequency TEPID data") +
        theme_bw() +
        ggtitle("Basho TE allele frequencies")

p
```

Save plot

```{r}
ggsave("../Plots/basho_alleles.pdf", p, height = 10, width = 15, units = "cm", useDingbats = FALSE)
```

```{r}
model <- lm(dat$Occ..Freq. ~ dat$Count_TE_present)
summary(model)
```